Using Refactoring Techniques to Reduce Duplication in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-04-29 sattose/paper.tex
Spelling & Formatting

\begin{abstract}
Duplication in source code can have a major negative impact on the maintainability of source code. There are several techniques that can be used in order to merge clones, reduce duplication, improve the design of the code and potentially also reduce the total volume of a software system. In this study, we look into the opportunities to aid in the process of refactoring these duplication problems for object-oriented programming languages. Measurements that have been conducted so far have indicated that more than half of the duplication in code is related to each other through inheritance, making it easier to refactor these clones in a clean way. Approximately a fifth of the duplication can be refactored through method extraction, the other clones require other techniques to be applied. More measurements will be conducted to get a detailed overview of in what contexts clones occur, and what this means for the refactoring processes of these clones. As a desired output, we strive to construct a model that automatically applies refactorings for a large part of the detected duplication problems. This model will be implemented for the Java programming language.
\end{abstract}

@@

Refactoring is used to improve quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler1999refactoring, wake2004refactoring}, that are integrated into most modern IDE's. However, most of these methods still require manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this research we will look at to what extend code clones can be automatically refactored.

Code clones can be found anywhere in a codebase. The location of a clone in the code has an impact on how it has to be refactored. Because of this, we will first look at where clones can be found. Using this information we can determine in which locations clones are found the most, on basis of which the refactoring will be prioritized. Apart from that, we will look at different types of clones. A duplicate part in a codebase does not always have to be an exact match with another part to be called a clone. We will look at the definitions of different types of clones and what the opportunities are to refactor duplicate parts of code if it's not an exact match. All of these measurements are performed on a large corpus of open source projects. We lay the main focus on the Java programming language as refactoring opportunities feature paradigm and programming language dependent aspects \cite{choi2011extracting}. However, most practises featured in this research will also be applicable with other object-oriented languages, like C\#.

\subsection{Background}
As code clones are seen as one of the most harmful types of technical debt, they have been studied very extensively. A survey by Roy et al \cite{roy2007survey} states definitions for various important concepts in code clone research. In this survey he mentions the concept ``clone pair'', which is \textit{a set of two code portions/fragments which are identical or similar to each other}. Furthermore, he defined ``clone class'' as \textit{the union of all clone pairs}. Apart from this, we use the definition ``clone instance'', which is a single code portion/fragment that is part of either a clone pair or clone class.

A tool called DCRA (Duplicated Code Refactoring Advisor) looks into refactoring opportunities for clone pairs \cite{fontana2012duplicated}. This tool chooses to only focus on refactoring clone pairs because they argue that \textit{clone pairs are much easier to manage when considered singularly.} As intermidiate steps the authors measure a corpus of Java systems for some clone-related properties of the systems, like the relation (in terms of inheritance) between code fragments in a clone pair.

Another tool called Aries \cite{higo2004aries, higo2008metric} focuses on the detection of refactoring oriented clones. In this paper the authors describe a few variables by which a clone can be assessed for refactoring suitability. This tool looks however only at the possibility to merge clones, rather than whether clones should be merged. Also, this tool only suggests refactorings, and does not apply them.

There are many other publications that look into code clone refactoring \cite{alwaqfi2017refactoring, chen2018clone, koni2001scenario}. However, none of the publications describing a refactoring tool have published their source code or application for refactoring support. Furthermore, all of these tools only support a subset of all harmful clones that are found. Also, these publications/tools are limited to suggesting refactoring opportunities for clones found, rather than actually applying refactorings where suitable. Finally, all published approaches have limitations, for instance false positives when detecting clones \cite{chen2018clone} or being limited to clone pairs \cite{higo2008metric}.

@@

\subsection{CloneRefactor}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. For our tool we decided to combine AST and Graph based approaches for clone detection, similar to Scorpio (which is a clone detection tool that's part of TinyPDG: a library for building intraprocedural program dependency graphs for Java programs). We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is useable with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. This way we build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of it's duplicates (verically). This is displayed in figure \ref{fig:clonerefactor}.

@@

\subsection{Thresholds}
CloneRefactor works on basis of three thresholds for finding clones:
\begin{enumerate}
  \item \textbf{Number of statements/declarations:} The amount of statements/declarations that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of tokens:} The amount of tokens (excluding whitespace, end-of-line terminators and comments) that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of lines:} The minimum amount of lines (excluding lines that do not contain any tokens, for tokens same exclusions apply) that should be equal/similar for it to be considered a clone.
\end{enumerate}
Of course, if we put any of these thresholds to zero it won't be taken into account anymore, so not all thresholds have to be used at all times. We consider ``number of lines'' to be the least important, as it highly depends on the programmer of the codebase (and we do not want this kind of dependence!). On basis of manual assessment, we have determined that setting the ``number of statements/declarations'' to 6 ensures that most non-harmful clones are filtered out. On the downside, this also filters out some harmful antipatterns, for instance if a cloned line has many tokens we might want to consider a clone if it spans less than 6 statements (as cloned line with many tokens is more harmful than one with few).

@@

\section{Code Clone Context}
To be able to refactor code clones, it is very important to consider the context of the clone. We define the following aspects of the clone as its context:
\begin{enumerate}
  \item The relation of clone instances among each other (for example: two clone instances in a clone class are part of the same object).
  \item Where a clone instance occurs in the code (for instance: a method-level clone is a clone instance that is (a part of) a single method).
  \item The contents of a clone instance (for instance: the clone instance consists of a one method declaration, a foreach statement and two variable declarations).
\end{enumerate}
Everything in the context of a clone has a big impact on how it has to be refactored. For this study we performed measurements on the context of clones in a large corpus of open source projects.

\subsection{The corpus}
For our measurements we use a large corpus of open source projects \cite{githubCorpus2013}\footnote{The corpus can be downloaded from the following URL: \url{http://groups.inf.ed.ac.uk/cup/javaGithub/java_projects.tar.gz}}. This corpus has been assembled to contain relatively higher quality projects (by filtering by forks). Also, any duplicate projects were removed from this corpus. This results in a variaty of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.

We then filtered the corpus further to make sure we are not including any test classes or generated classes. Many Java/Maven projects use a structure where they separate the application and it's tests in the different folders (``/src/main/java'' and ``/src/test/java'' respectively). Because of this, we chose to only use projects from the corpus which use this structure (and had at least a ``/src/main/java'' folder). To limit the execution time of the script, we also decided to limit the maximum amount of source files in a single project to 1.000 (projects with more source files were not considered, which filtered only 5 extra projects out of the corpus). Of the 14.436 projects in the corpus over 3.848 remained, which is plenty for our purposes. The script to filter the corpus in included in our GitHub repository \footnote{The script we use to filter the corpus: \url{https://github.com/SimonBaars/CloneRefactor/blob/MeasurementsVersion1/src/main/java/com/simonbaars/clonerefactor/scripts/PrepareProjectsFolder.java}}.

@@

\begin{table}[H]
  \begin{center}
  \caption{CloneRefactor results for Java projects corpus \cite{githubCorpus2013}.} \label{table:general}
  \bigskip
\begin{tabular}{|l|l|}
\hline
Amount of projects                                                                                      & 3.848      \\ \hline
Amount of lines                                                                                         & 8.284.140  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines (excluding\\ whitespace, comments, etc.)\end{tabular}        & 8.163.429  \\ \hline
Amount of statements/declarations                                                                       & 6.863.725  \\ \hline
Amount of tokens                                                                                        & 66.964.270 \\ \hline
Amount of lines cloned                                                                                  & 1.341.094  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines cloned (excluding\\ whitespace, comments, etc.)\end{tabular} & 815.799    \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of statements/declarations\\ cloned\end{tabular}                      & 747.993    \\ \hline
Amount of tokens cloned                                                                                 & 9.800.819  \\ \hline
Amount of clone classes                                                                                 & 34.367     \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

The ranking of the previous list of categories also matters, as it shows the different levels in which clones were assessed. For instance, if two clone instances of a clone class belong to the ``same method'' category but the third belongs to the ``same class'', we will always chose the item lowest on the list.

@@

\begin{table}[H]
  \begin{center}
  \caption{Clone relations} \label{table:relations}
  \bigskip
\begin{tabular}{|l|l|l|} \hline
\textbf{Relation}  & \textbf{Amount} & \textbf{Percentage} \\ \hline
Unrelated          & 13.529           & 39,37               \\ \hline
Same Class          & 8.341            & 24,28               \\ \hline
Sibling            & 5.978            & 17,40               \\ \hline
Same Method         & 2.456            & 7,15                \\ \hline
External Superclass & 2.402            & 6,99                \\ \hline
First Cousin        & 695             & 2,02                \\ \hline
Superclass         & 489             & 1,42                \\ \hline
Common Hierarchy    & 442             & 1,29                \\ \hline
Ancestor           & 28              & 0,08               \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\subsection{Clone instance location}
After mapping the relations between individual clones, we looked at the location of individual clone instances. A paper by Lozano et al \cite{lozano2007evaluating} discusses the harmfulness of cloning. In this paper the author argues that 98\% are produced at method-level. However, this claim is based on a small dataset and based on human copy-paste behaviour rather than static code analysis. We validated this claim over the corpus we use. For this, we chose the following categories:
\begin{enumerate}
  \item \textbf{Method/Constructor Level:} A clone instance that does not exceed the boundaries of a single method or constructor (optionally including the declaration of the method or constructor itself).
  \item \textbf{Class Level:} A clone instance in a class, that exceeds the boundaries of a single method or contains something else in the class (like field declarations, other methods, etc.).
  \item \textbf{Interface Level:} A clone that is (a part of) an interface.
  \item \textbf{Enumeration Level:} A clone that is (a part of) an enumeration.
\end{enumerate}

@@

\begin{table}[H]
  \begin{center}
  \caption{Clone instance locations} \label{table:locations}
  \bigskip
\begin{tabular}{|l|l|l|}
\hline
\textbf{Location}  & \textbf{Amount}  & \textbf{Percentage} \\ \hline
Method Level       & 19.818           & 57,68               \\ \hline
Class Level        & 13.259           & 38,59               \\ \hline
Constructor Level  & 1.099            & 3,20                \\ \hline
Interface Level    & 110              & 0,32                \\ \hline
Enum Level         & 74               & 0,22                \\ \hline
\end{tabular}
\end{center}
\end{table}

From these results we can see that the claim of 98\% of clones being produced at method-level is nowhere near correct. In fact, around 58\% of the clones are produced at method-level. About 39\% of clones either span several methods/constructors or contain something like a field declaration. Another 3\% of the clones is found in contructors. The amount of clones found in interfaces and enumerations is very low.

\subsection{Clone instance contents}
Finally, we looked at the contents of individual clone instances: what kind of declarations and statements do they span. We selected the following categories to be relevant for refactoring:
\begin{enumerate}
  \item \textbf{Full Method/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans a method partially, optionally including its declaration.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods (so not fields or anything in between).
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Includes Fields/Constructor:} A clone that spans a combination of fields and other things, like methods.
  \item \textbf{Method/Class/Interface/Enumeration Declaration:} A clone that contains the declaration (usually the first line) of a class, method, interface or enumeration.
  \item \textbf{Other:} Anything that does not match with above stated categories.
\end{enumerate}

The results for these categories are displayed are displayed in table \ref{table:contents}.

\begin{table}[H]
  \begin{center}
  \caption{Clone instance contents} \label{table:contents}
  \bigskip
\begin{tabular}{|l|l|l|}
  \hline
  \textbf{Contents}                                                        & \textbf{Amount} & \textbf{Percentage} \\ \hline
  Partial Method                                                           & 19.264          & 56,07               \\ \hline
  Several Methods                                                          & 9.076           & 26,41               \\ \hline
  Includes Constructor                                                     & 1.528           & 4,45                \\ \hline
  Includes Field                                                           & 1.149           & 3,34                \\ \hline
  Partial Constructor                                                      & 1.098           & 3,20                \\ \hline
  Only Fields                                                              & 565             & 1,64                \\ \hline
  Full Method                                                              & 554             & 1,61                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Class\\ Declaration\end{tabular}     & 445             & 1,30                \\ \hline
  Other                                                                    & 369             & 1,07                \\ \hline
  Full Class                                                               & 192             & 0,56                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Constant\end{tabular}         & 52              & 0,15                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Declaration\end{tabular}      & 47              & 0,14                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Interface\\ Declaration\end{tabular} & 10              & 0,03                \\ \hline
  Full Interface                                                           & 6               & 0,02                \\ \hline
  Full Enumeration                                                         & 4               & 0,01                \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\textbf{Type 1:} Identical code fragments except for variations in whitespace (may be also variations in layout) and comments.\\
\textbf{Type 2:} Structurally/syntactically identical fragments except for variations in identifiers, literals, types, layout and comments.\\
\textbf{Type 3:} Copied fragments with further modifications. Statements can be changed, added or removed in addition to variations in identifiers, literals, types, layout and comments.\\
\textbf{Type 4:} Two or more code fragments that perform the same computation but implemented through different syntactic variants.

A higher type of clone means that it's harder to detect. It also makes the clone harder to refactor, as more transformations would be required. Higher clone types also become more disputable whether they actually indicate a harmful antipattern (not every clone is harmful). Type 1 clones are most often, if not always with the right thresholds, harmful. Type 2 and type 3 clones are more disputable whether they are harmful and whether they can be merged in a way that improves the design of the code. In this chapter we will redefine type 2 and type 3 clones to increase the chance that we can merge the clone while improving the design. Type-4 clones are, because of the serious challenges involved in their detection and refactoring, out of the scope for this study.

\subsection{Type 2 clones}
Type 2 clones, by definition, allow any change in identifiers, literals, types, layout and comments. For refactoring purposes, this doesn't always make a lot of sense. If we allow any change in identifiers, literals and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones. Merging such clones, if possible at all, might only proof to be harmful.

Because of that, for this research we have looked into a redefinition of type 2 clones to be able to detect such clones that can and should be merged. Our definition ensures functional similarity by applying the following changes to type 2 clones:

\begin{itemize}
  \item \textbf{Considering types:} The definition of type 2 clones defines that types should not be considered. We disagree, because types can make a significant change to the meaning of a code segment and thus whether this segment should be considered a clone. For instance, we should not say that an Apple type object is the same as an object of Orange type (don't compare apples with oranges).
  \item \textbf{Having a distinction between different variables:} By the definition of type 2 clones any identifiers would not be taken into account. We agree that a difference in identifiers may still result in a harmful clone, but we should still consider the distinction between different variables. For instance, if we call a method like this: \textit{``myMethod(var1, var2)''}, or call this method like this: \textit{``myMethod(var1, var1)''}. Even if the variables have the same type, the distinction between the variables is important to ensure the functionality is the same after merging.
  \item \textbf{Defining a threshold for variability in literals:} By the definition of type 2 clones any literals would not be taken into account. We agree, as when merging the clone (for instance by extracting a method), we can simply turn the literal into a method parameter. However, we would argue that thresholds matter here. How many literals may differ for the segment still to be considered a clone with another segment? We need to define a threshold to be sure that, by merging, we are not replacing a code fragment by a worse maintainable design.
\end{itemize}

@@

However, not always can a clone easily be merged. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, cloned code can contain a complex control structure, like labels, return, break, continue, etc. In such instances, more conditions apply do be able to conduct a refactoring, if advisable at all.

The most trivial way to merge a clone is though method extraction. However, method extraction is not always possible. For instance, if a part of a subtree (in a programs' AST) is matched as a clone. Because of this, we chose to measure what percentage of ``Partial Method'' clones are refactorable using method extraction. Our results are displayed in table \ref{table:refactorability}.

\begin{table}[H]
  \begin{center}
  \caption{Clone instance contents} \label{table:refactorability}
  \bigskip
\begin{tabular}{|l|l|l|}
\hline
\textbf{Refactorability}        & \textbf{Amount} & \textbf{Percentage} \\ \hline
Cannot directly be extracted    & 10990           & 41,91               \\ \hline
Can be extracted                & 5791            & 22,08               \\ \hline
Not possible because of content & 9444            & 36,01               \\ \hline
\end{tabular}
\end{center}
\end{table}

>>>>

\begin{abstract}
Duplication in source code can have a major negative impact on the maintainability of source code. There are several techniques that can be used in order to merge clones, reduce duplication, improve the design of the code and potentially also reduce the total volume of a software system. In this study, we look into the opportunities to aid in the process of refactoring these duplication problems for object-oriented programming languages. Measurements that have been conducted so far have indicated that more than half of the duplication in code is related to each other through inheritance, making it easier to refactor these clones in a clean way. Approximately a fifth of the duplication can be refactored through method extraction, the other clones require other techniques to be applied. More measurements will be conducted to get a detailed overview of in what contexts clones occur, and what this means for the refactoring processes of these clones. As the desired output, we strive to construct a model that automatically applies refactorings for a large part of the detected duplication problems. This model will be implemented for the Java programming language.
\end{abstract}

@@

Refactoring is used to improve the quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler1999refactoring, wake2004refactoring}, that are integrated into most modern IDE's. However, most of these methods still require a manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this research, we will look at to what extend code clones can be automatically refactored.

Code clones can be found anywhere in a codebase. The location of a clone in the code has an impact on how it has to be refactored. Because of this, we will first look at where clones can be found. Using this information we can determine in which locations clones are found the most, on basis of which the refactoring will be prioritized. Apart from that, we will look at different types of clones. A duplicate part in a codebase does not always have to be an exact match with another part to be called a clone. We will look at the definitions of different types of clones and what the opportunities are to refactor duplicate parts of code if it's not an exact match. All of these measurements are performed on a large corpus of open source projects. We lay the main focus on the Java programming language as refactoring opportunities feature paradigm and programming language dependent aspects \cite{choi2011extracting}. However, most practices featured in this research will also be applicable to other object-oriented languages, like C\#.

\subsection{Background}
As code clones are seen as one of the most harmful types of technical debt, they have been studied very extensively. A survey by Roy et al \cite{roy2007survey} states definitions for various important concepts in code clone research. In this survey, he mentions the concept ``clone pair'', which is \textit{a set of two code portions/fragments which are identical or similar to each other}. Furthermore, he defined ``clone class'' as \textit{the union of all clone pairs}. Apart from this, we use the definition ``clone instance'', which is a single code portion/fragment that is part of either a clone pair or clone class.

A tool called DCRA (Duplicated Code Refactoring Advisor) looks into refactoring opportunities for clone pairs \cite{fontana2012duplicated}. This tool chooses to only focus on refactoring clone pairs because they argue that \textit{clone pairs are much easier to manage when considered singularly.} As intermediate steps, the authors measure a corpus of Java systems for some clone-related properties of the systems, like the relation (in terms of inheritance) between code fragments in a clone pair.

Another tool called Aries \cite{higo2004aries, higo2008metric} focuses on the detection of refactoring oriented clones. In this paper, the authors describe a few variables by which a clone can be assessed for refactoring suitability. This tool looks however only at the possibility to merge clones, rather than whether clones should be merged. Also, this tool only suggests refactorings and does not apply them.

There are many other publications that look into code clone refactoring \cite{alwaqfi2017refactoring, chen2018clone, koni2001scenario}. However, none of the publications describing a refactoring tool have published their source code or application for refactoring support. Furthermore, all of these tools only support a subset of all harmful clones that are found. Also, these publications/tools are limited to suggesting refactoring opportunities for clones found, rather than actually applying refactorings where suitable. Finally, all published approaches have limitations, for instance, false positives when detecting clones \cite{chen2018clone} or being limited to clone pairs \cite{higo2008metric}.

@@

\subsection{CloneRefactor}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. For our tool, we decided to combine AST- and Graph-based approaches for clone detection, similar to Scorpio (which is a clone detection tool that's part of TinyPDG: a library for building intraprocedural program dependency graphs for Java programs). We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is useable with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. This way we build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of its duplicates (vertically). This is displayed in figure \ref{fig:clonerefactor}.

@@

\subsection{Thresholds}
CloneRefactor works on basis of three thresholds for finding clones:
\begin{enumerate}
  \item \textbf{Number of statements/declarations:} The number of statements/declarations that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of tokens:} The number of tokens (excluding whitespace, end-of-line terminators and comments) that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of lines:} The minimum amount of lines (excluding lines that do not contain any tokens, for tokens same exclusions apply) that should be equal/similar for it to be considered a clone.
\end{enumerate}
Of course, if we put any of these thresholds to zero it won't be taken into account anymore, so not all thresholds have to be used at all times. We consider ``number of lines'' to be the least important, as it highly depends on the programmer of the codebase (and we do not want this kind of dependence!). On basis of manual assessment, we have determined that setting the ``number of statements/declarations'' to 6 ensures that most non-harmful clones are filtered out. On the downside, this also filters out some harmful antipatterns, for instance, if a cloned line has many tokens we might want to consider a clone if it spans less than 6 statements (as a cloned line with many tokens is more harmful than one with few).

@@

\section{Code Clone Context}
To be able to refactor code clones, it is very important to consider the context of the clone. We define the following aspects of the clone as its context:
\begin{enumerate}
  \item The relation of clone instances among each other (for example two clone instances in a clone class are part of the same object).
  \item Where a clone instance occurs in the code (for instance: a method-level clone is a clone instance that is (a part of) a single method).
  \item The contents of a clone instance (for instance: the clone instance consists of a one method declaration, a foreach statement, and two variable declarations).
\end{enumerate}
Everything in the context of a clone has a big impact on how it has to be refactored. For this study, we performed measurements on the context of clones in a large corpus of open source projects.

\subsection{The corpus}
For our measurements we use a large corpus of open source projects \cite{githubCorpus2013}\footnote{The corpus can be downloaded from the following URL: \url{http://groups.inf.ed.ac.uk/cup/javaGithub/java_projects.tar.gz}}. This corpus has been assembled to contain relatively higher quality projects (by filtering by forks). Also, any duplicate projects were removed from this corpus. This results in a variety of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.

We then filtered the corpus further to make sure we are not including any test classes or generated classes. Many Java/Maven projects use a structure where they separate the application and its tests in the different folders (``/src/main/java'' and ``/src/test/java'' respectively). Because of this, we chose to only use projects from the corpus which use this structure (and had at least a ``/src/main/java'' folder). To limit the execution time of the script, we also decided to limit the maximum amount of source files in a single project to 1.000 (projects with more source files were not considered, which filtered only 5 extra projects out of the corpus). Of the 14.436 projects in the corpus over 3.848 remained, which is plenty for our purposes. The script to filter the corpus in included in our GitHub repository \footnote{The script we use to filter the corpus: \url{https://github.com/SimonBaars/CloneRefactor/blob/MeasurementsVersion1/src/main/java/com/simonbaars/clonerefactor/scripts/PrepareProjectsFolder.java}}.

@@

\begin{table}[H]
  \begin{center}
  \caption{CloneRefactor results for Java projects corpus \cite{githubCorpus2013}.} \label{table:general}
  \bigskip
\begin{tabular}{|l|l|}
\hline
Amount of projects                                                                                      & 3,848      \\ \hline
Amount of lines                                                                                         & 8,284,140  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines (excluding\\ whitespace, comments, etc.)\end{tabular}        & 8,163,429  \\ \hline
Amount of statements/declarations                                                                       & 6,863,725  \\ \hline
Amount of tokens                                                                                        & 66,964,270 \\ \hline
Amount of lines cloned                                                                                  & 1,341,094  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines cloned (excluding\\ whitespace, comments, etc.)\end{tabular} & 815,799    \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of statements/declarations\\ cloned\end{tabular}                      & 747,993    \\ \hline
Amount of tokens cloned                                                                                 & 9,800,819  \\ \hline
Amount of clone classes                                                                                 & 34,367     \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

The ranking of the previous list of categories also matters as it shows the different levels in which clones were assessed. For instance, if two clone instances of a clone class belong to the ``same method'' category but the third belongs to the ``same class'', we will always choose the item lowest on the list.

@@

\begin{table}[H]
  \begin{center}
  \caption{Clone relations} \label{table:relations}
  \bigskip
\begin{tabular}{|l|l|l|} \hline
\textbf{Relation}  & \textbf{Amount} & \textbf{Percentage} \\ \hline
Unrelated          & 13,529           & 39.37               \\ \hline
Same Class          & 8,341            & 24.28               \\ \hline
Sibling            & 5,978            & 17.40               \\ \hline
Same Method         & 2,456            & 7.15                \\ \hline
External Superclass & 2,402            & 6.99                \\ \hline
First Cousin        & 695             & 2.02                \\ \hline
Superclass         & 489             & 1.42                \\ \hline
Common Hierarchy    & 442             & 1.29                \\ \hline
Ancestor           & 28              & 0.08               \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\subsection{Clone instance location}
After mapping the relations between individual clones, we looked at the location of individual clone instances. A paper by Lozano et al \cite{lozano2007evaluating} discusses the harmfulness of cloning. In this paper, the author argues that 98\% are produced at method-level. However, this claim is based on a small dataset and based on human copy-paste behavior rather than static code analysis. We validated this claim over the corpus we use. For this, we chose the following categories:
\begin{enumerate}
  \item \textbf{Method/Constructor Level:} A clone instance that does not exceed the boundaries of a single method or constructor (optionally including the declaration of the method or constructor itself).
  \item \textbf{Class Level:} A clone instance in a class, that exceeds the boundaries of a single method or contains something else in the class (like field declarations, other methods, etc.).
  \item \textbf{Interface Level:} A clone that is (a part of) an interface.
  \item \textbf{Enumeration Level:} A clone that is (a part of) an enumeration.
\end{enumerate}

@@

\begin{table}[H]
  \begin{center}
  \caption{Clone instance locations} \label{table:locations}
  \bigskip
\begin{tabular}{|l|l|l|}
\hline
\textbf{Location}  & \textbf{Amount}  & \textbf{Percentage} \\ \hline
Method Level       & 19,818           & 57.68               \\ \hline
Class Level        & 13,259           & 38.59               \\ \hline
Constructor Level  & 1,099            & 3.20                \\ \hline
Interface Level    & 110              & 0.32                \\ \hline
Enum Level         & 74               & 0.22                \\ \hline
\end{tabular}
\end{center}
\end{table}

From these results, we can see that the claim of 98\% of clones being produced at method-level is nowhere near correct. In fact, around 58\% of the clones are produced at method-level. About 39\% of clones either span several methods/constructors or contain something like a field declaration. Another 3\% of the clones are found in constructors. The amount of clones found in interfaces and enumerations is very low.

\subsection{Clone instance contents}
Finally, we looked at the contents of individual clone instances: what kind of declarations and statements do they span. We selected the following categories to be relevant for refactoring:
\begin{enumerate}
  \item \textbf{Full Method/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans a method partially, optionally including its declaration.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods (so not fields or anything in between).
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Includes Fields/Constructor:} A clone that spans a combination of fields and other things, like methods.
  \item \textbf{Method/Class/Interface/Enumeration Declaration:} A clone that contains the declaration (usually the first line) of a class, method, interface or enumeration.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

The results for these categories are displayed in table \ref{table:contents}.

\begin{table}[H]
  \begin{center}
  \caption{Clone instance contents} \label{table:contents}
  \bigskip
\begin{tabular}{|l|l|l|}
  \hline
  \textbf{Contents}                                                        & \textbf{Amount} & \textbf{Percentage} \\ \hline
  Partial Method                                                           & 19,264          & 56.07               \\ \hline
  Several Methods                                                          & 9,076           & 26.41               \\ \hline
  Includes Constructor                                                     & 1,528           & 4.45                \\ \hline
  Includes Field                                                           & 1,149           & 3.34                \\ \hline
  Partial Constructor                                                      & 1,098           & 3.20                \\ \hline
  Only Fields                                                              & 565             & 1.64                \\ \hline
  Full Method                                                              & 554             & 1.61                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Class\\ Declaration\end{tabular}     & 445             & 1.30                \\ \hline
  Other                                                                    & 369             & 1.07                \\ \hline
  Full Class                                                               & 192             & 0.56                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Constant\end{tabular}         & 52              & 0.15                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Declaration\end{tabular}      & 47              & 0.14                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Interface\\ Declaration\end{tabular} & 10              & 0.03                \\ \hline
  Full Interface                                                           & 6               & 0.02                \\ \hline
  Full Enumeration                                                         & 4               & 0.01                \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\textbf{Type 1:} Identical code fragments except for variations in whitespace (may be also variations in layout) and comments.\\
\textbf{Type 2:} Structurally/syntactically identical fragments except for variations in identifiers, literals, types, layout and comments.\\
\textbf{Type 3:} Copied fragments with further modifications. Statements can be changed, added or removed in addition to variations in identifiers, literals, types, layout, and comments.\\
\textbf{Type 4:} Two or more code fragments that perform the same computation but implemented through different syntactic variants.

A higher type of clone means that it's harder to detect. It also makes the clone harder to refactor, as more transformations would be required. Higher clone types also become more disputable whether they actually indicate a harmful antipattern (not every clone is harmful). Type 1 clones are most often, if not always with the right thresholds, harmful. Type 2 and type 3 clones are more disputable whether they are harmful and whether they can be merged in a way that improves the design of the code. In this chapter, we will redefine type 2 and type 3 clones to increase the chance that we can merge the clone while improving the design. Type-4 clones are, because of the serious challenges involved in their detection and refactoring, out of the scope for this study.

\subsection{Type 2 clones}
Type 2 clones, by definition, allow any change in identifiers, literals, types, layout, and comments. For refactoring purposes, this doesn't always make a lot of sense. If we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones. Merging such clones, if possible at all, might only prove to be harmful.

Because of that, for this research, we have looked into a redefinition of type 2 clones to be able to detect such clones that can and should be merged. Our definition ensures functional similarity by applying the following changes to type 2 clones:

\begin{itemize}
  \item \textbf{Considering types:} The definition of type 2 clones states that types should not be considered. We disagree because types can make a significant change to the meaning of a code segment and thus whether this segment should be considered a clone. For instance, we should not say that an Apple type object is the same as an object of Orange type (don't compare apples with oranges).
  \item \textbf{Having a distinction between different variables:} By the definition of type 2 clones, any identifiers would not be taken into account. We agree that a difference in identifiers may still result in a harmful clone, but we should still consider the distinction between different variables. For instance, if we call a method like this: \textit{``myMethod(var1, var2)''}, or call this method like this: \textit{``myMethod(var1, var1)''}. Even if the variables have the same type, the distinction between the variables is important to ensure the functionality is the same after merging.
  \item \textbf{Defining a threshold for variability in literals:} By the definition of type 2 clones any literals would not be taken into account. We agree, as when merging the clone (for instance by extracting a method), we can simply turn the literal into a method parameter. However, we would argue that thresholds matter here. How many literals may differ for the segment still to be considered a clone with another segment? We need to define a threshold to be sure that, by merging, we are not replacing a code fragment by a worse maintainable design.
\end{itemize}

@@

However, not always can a clone easily be merged. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, cloned code can contain a complex control structure, like labels, return, break, continue, etc. In such instances, more conditions apply to be able to conduct a refactoring, if advisable at all.

The most trivial way to merge a clone is through method extraction. However, method extraction is not always possible. For instance, if a part of a subtree (in a programs' AST) is matched as a clone. Because of this, we chose to measure what percentage of ``Partial Method'' clones are refactorable using method extraction. Our results are displayed in table \ref{table:refactorability}.

\begin{table}[H]
  \begin{center}
  \caption{Clone instance contents} \label{table:refactorability}
  \bigskip
\begin{tabular}{|l|l|l|}
\hline
\textbf{Refactorability}        & \textbf{Amount} & \textbf{Percentage} \\ \hline
Cannot directly be extracted    & 10,990           & 41.91               \\ \hline
Can be extracted                & 5,791            & 22.08               \\ \hline
Not possible because of content & 9,444            & 36.01               \\ \hline
\end{tabular}
\end{center}
\end{table}

From this table, we can see that approximately 22 percent of the clones can be refactored through method extraction. For the other clones, we must first build a catalog of appropriate refactorings.
